Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-19 paper/main.tex
write introductory note for 2nd application; link to previous chapters

% --------------------------------------------------------------------------------------------------
\subsection{Position Extraction}

@@

\subsubsection{Imaging-based Comparison Algorithm}
The preceding algorithm is performant and effective but can not be applied in the event of a
division process.
Due to possibly differing division times, a cell which is present in the simulation might not yet
have a corresponding cell in the experimental data at this time point yet (or vice-versa) (see
Figure~\ref{fig:mask-difference-metric} (E)).
In order to still be able to compare these datapoints, we require additional methods which can be
used for comparisons across division events.
\textbf{[Es bleibt unklar, was du hier verhandeln möchtest.]}
\textit{Jetzt klarer?}

@@

In this section we apply the previously developed methods to the problem of parameter estimation
of our model.
We cover 3 distinct cases with increasing level of complexity.
In the first study, we utilize results previously obtained by \cite{Amir2014,Amir2014_2} who
investigated elastic deformations of \textit{E.Coli} that are partially trapped within a narrow
tube and subject to a force resulting from a fluidic flow.
This setup does not involve any cell-cell interactions but specifically requires a flexible rod.
It is therefore well suited as a starting point to assess the usefulness of our methods and the
ability of our model to describe these results.\\
The last two cases consider time-series of microscopic images containing multiple cells (see
subsection~\ref{section:application-2}) which will be used to perform a parameter estimation of our
full mechanical model.
The approach is split into a first part which considers a time interval that does not contain any
division events and a second part which extends the initially chosen interval to include division
events.
In the first part, we utilize the direct comparison algorithm while the latter requires usage of the
imaging-based comparison algorithm.

@@

%###################################################################################################
\subsection{Application 2: \textit{E.Coli.} growing on Agar}
\label{section:application-2}
To show the usefulness of our previously developed methods
(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of
individually growing \textit{E.Coli} using a sequence of microscopic images.
The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
section~\ref{section:data-code-availability}) has been automatically segmented with the
omnipose~\cite{Cutler2022} algorithm to obtain the required cell masks.

@@

>>>>

% --------------------------------------------------------------------------------------------------
\subsection{Position Extraction}
\label{subsec:position-extraction-algorithm}

@@

\subsubsection{Imaging-based Comparison Algorithm}
\label{subsec:imaging-based-comparison}
The preceding algorithm is performant and effective but can not be applied in the event of a
division process.
Due to possibly differing division times, a cell which is present in the simulation might not yet
have a corresponding cell in the experimental data at this time point yet (or vice-versa) (see
Figure~\ref{fig:mask-difference-metric} (E)).
In order to still be able to compare these datapoints, we require additional methods which can be
used for comparisons across division events.
\textbf{[Es bleibt unklar, was du hier verhandeln möchtest.]}
\textit{Jetzt klarer?}

@@

In this section we apply the previously developed methods to the problem of parameter estimation
of our model.
We cover 3 distinct cases with increasing level of complexity.
In the first study, we utilize results previously obtained by Amir et. al~\cite{Amir2014,Amir2014_2}
who investigated elastic deformations of \textit{E.Coli} that are partially trapped within a narrow
tube and subject to a force resulting from a fluidic flow.
This setup does not involve any cell-cell interactions but specifically requires a flexible rod.
It is therefore well suited as a starting point to assess the usefulness of our methods and the
ability of our model to describe these results.\\
The last two cases consider time-series of microscopic images containing multiple cells (see
subsection~\ref{section:application-2}) which will be used to perform a parameter estimation of our
full mechanical model.
The approach is split into a first part which considers a time interval that does not contain any
division events and a second part which extends the initially chosen interval to include division
events.
In the first part, we cover a time interval which does not contain division events and thus utilize
the direct comparison algorithm while the second case contains division events and in turn requires
usage of the imaging-based comparison algorithm.

@@

%###################################################################################################
\subsection{Application 2: \textit{E.Coli.} growing on Agar}
\label{section:application-2}

In this section we study a collection of growing, dividing and physically interacting
\textit{E.Coli}.
We use a series of microscopic images and corresponding segmentation masks for the individual
bacteria as the underlying dataset which is partially displayed in
Figure~\ref{fig:application-2-data-overview}.
With the previously developed extraction algorithm (see
section~\ref{subsec:position-extraction-algorithm}) we obtain the positions of our agents at all
time points.
These can be used as initial values and later within our parameter estimation routine.
We initialize our computational model (see section~\ref{sec:computational-model}) with the obtained
positions and optimize the parameters in order to find a set which best describes the observed
results.
The first case considers a time interval without cell-division while the second case does contain
division events.
Therefore, the first approach will optimize the direct comparison cost function (see
section~\ref{subsec:direct-comparison}) while the latter approach has to account for division events
and thus requires the imaging-based method (see section~\ref{subsec:imaging-based-comparison}).
These two scenarios present an opportunity to apply the previously developed methods within a
realistic context and validate our approach.

To show the usefulness of our previously developed methods
(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of
individually growing \textit{E.Coli} using a sequence of microscopic images.
The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
section~\ref{section:data-code-availability}) has been automatically segmented with the
omnipose~\cite{Cutler2022} algorithm to obtain the required cell masks.

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001044.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001054.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001063.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001074.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \caption{TODO insert segmentation masks here}
    \label{fig:application-2-data-overview}
\end{figure}
